---
title: "Take-home_Ex2"
author: "CHEN HO TUNG"
date: "March 5, 2026"
date-modified: "last-modified"

format:
  html:
    toc-depth: 4
    toc-expand: true

execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# **Take-home Exercise 2: Prototyping Modules for Visual Analytics Shiny Application**

# 2. Fintech Customer Behavior and Risk Analytics

## 2.1 Overview

Behavioral and trancsactional data from 48723 customers of a Colombian fintech company collected over 12 months from January 4, 2023, to December 29, 2023. Comprises 3,159,157 individual transactions.

## 2.2 Setting the Scene

To analyse temporal patterns in customer financial behavior and engagement using the given COFINFAD Colombian Fintech Financial Analytics Dataset with the aim of identifying the key factors that influence customer lifetime value (CLV) and predicting the long-term financial value generated by customers. In the fintech industry, understanding which customer characteristics and behavioral patterns. The study aims to develop a predictive modelling framework to estimate Customer Lifetime Value (CLV) by identifying the key behavioral and transactional factors that influence long-term customer profitability. To achieve this objective, Decision Tree Regression and Random Forest Regression models are implemented to predict customer lifetime value based on customer financial behavior and engagement indicators, allowing the analysis to both generate accurate predictions and identify the most important drivers of customer value within the fintech platform.

## 2.3 The Task

In this take-home exercise, you are required to select one of the module of your proposed Shiny application and complete the following tasks:

-   To evaluate and determine the necessary R packages needed for your Shiny application are supported in **R CRAN**,

-   To prepare and test the specific R codes can be run and returned the correct visual output as expected,

-   To determine the parameters and outputs that will be exposed on the Shiny applications, and

-   To select the appropriate Shiny UI components for exposing the parameters determine above.

## 2.4 Getting Started

We load the following R packages using the `pacman::p_load()` function:

-   **dplyr**: Provides efficient functions for data manipulation such as filtering, selecting, grouping, and summarising variables.
-   **tidyverse**: A collection of essential packages for data science workflows including data wrangling, transformation, and visualization.
-   **janitor**: Helps clean and standardise variable names, making datasets easier to work with during analysis.
-   **readxl**: Allows importing data from Excel files into R.
-   **ggstatsplot**: Extends `ggplot2` by integrating statistical tests, effect sizes, and annotations directly into visualisations.
-   **ggrepel**: Prevents overlapping text labels in `ggplot2` graphics, improving readability of annotated plots.
-   **patchwork**: Enables the creation of composite visualisations by combining multiple `ggplot2` plots into a single layout.
-   **colorspace**, **ggthemes**, and **hrbrthemes**: Provide additional colour palettes and themes to enhance the visual appearance of plots.
-   **ggdist** and **ggridges**: Provide advanced visualisation tools for displaying distributions and density plots.
-   **FunnelPlotR**: Used to generate funnel plots for visualising variation and performance comparisons.
-   **gganimate** and **gifski**: Support animated visualisations that illustrate how patterns evolve over time.
-   **knitr**: Used for generating reproducible reports in R Markdown or Quarto.
-   **rpart** and **rpart.plot**: Used to build and visualise decision tree models for predicting customer lifetime value.
-   **visNetwork** and **visTree**: Provide interactive visualisations of decision trees and network structures.
-   **caret**: Provides a unified interface for machine learning workflows, including model training, validation, and preprocessing.
-   **ranger**: Implements an efficient and fast Random Forest algorithm suitable for large datasets.
-   **htmlwidgets**: Enables interactive visualisations to be exported as HTML components for viewing in browsers or Shiny applications.

#### Import Package

```{r}
pacman::p_load(dplyr, ggstatsplot, tidyverse,ggrepel, patchwork, janitor,colorspace,ggstatsplot,ggdist,ggridges,FunnelPlotR,knitr,gganimate,gifski,gapminder,ggthemes, hrbrthemes,tidyverse,readxl, rpart, visNetwork, visTree,rpart.plot, htmlwidgets, caret,ranger,patchwork,Metrics, reshape2)
```

## 2.5 Load Data

```{r}
customer_data <- read_csv("data/customer_data.csv")
customer_data
```

```{r}
transactions_data <- read_csv("data/transactions_data.csv")
transactions_data
```

## 2.6 Data Pre-processing

### 2.61 Glimpse of data

The customer dataset contains demographic, behavioural, engagement, satisfaction, and financial indicators for fintech users. The glimpse() function is used to inspect the structure of the dataset.

```{r}
glimpse(customer_data)
```

### 2.6.2 Check for duplication

By using the `duplicated()` function,

```{r}
duplicates <- customer_data[duplicated(customer_data), ]
duplicates
```

```{r}
customer_data$customer_id[duplicated(customer_data$customer_id)]
```

### 2.6.3 Filtering data for selected variables

For the Decision Tree analysis, relevant customer-level attributes are selected to build the modelling dataset.

```{r}
df_analysis <- customer_data[, c(
"age",
"household_size",
"credit_utilization_ratio",
"tx_count",
"avg_tx_value",
"total_tx_volume",
"transaction_frequency",
"monthly_transaction_count",
"average_transaction_value",
"weekend_transaction_ratio",
"avg_daily_transactions",
"customer_tenure",
"satisfaction_score",
"nps_score",
"app_store_rating",
"support_tickets_count",
"resolved_tickets_ratio",
"customer_lifetime_value"
)]
```

The selected variables are described below.

### 2.6.4 Check data structure

The structure of the filtered dataset is inspected using the `str()` function.

```{r}
str(df_analysis)
```

### 2.6.5 Re-casting variables

Re-cast variables for interpretability in later visualization.

#### 2.6.5.1 Re-cast numeric variable

```{r}
df_analysis$age <- as.numeric(df_analysis$age)

df_analysis$household_size <- as.numeric(df_analysis$household_size)

df_analysis$credit_utilization_ratio <- as.numeric(df_analysis$credit_utilization_ratio)

df_analysis$transaction_frequency <- as.numeric(df_analysis$transaction_frequency)

df_analysis$monthly_transaction_count <- as.numeric(df_analysis$monthly_transaction_count)

df_analysis$average_transaction_value <- as.numeric(df_analysis$average_transaction_value)

df_analysis$avg_daily_transactions <- as.numeric(df_analysis$avg_daily_transactions)

df_analysis$customer_tenure <- as.numeric(df_analysis$customer_tenure)

df_analysis$satisfaction_score <- as.numeric(df_analysis$satisfaction_score)

df_analysis$nps_score <- as.numeric(df_analysis$nps_score)

df_analysis$app_store_rating <- as.numeric(df_analysis$app_store_rating)

df_analysis$support_tickets_count <- as.numeric(df_analysis$support_tickets_count)

df_analysis$resolved_tickets_ratio <- as.numeric(df_analysis$resolved_tickets_ratio)

df_analysis$customer_lifetime_value <- as.numeric(df_analysis$customer_lifetime_value)
```

### 2.6.6 Check and Remove Missing Values

The missing values for each variable are checked using:

```{r}
colSums(is.na(df_analysis)) / nrow(df_analysis) * 100
```

```{r}
df_analysis <- na.omit(df_analysis)
```

#### Preview Pre-Processed Dataframe

```{r}
head(df_analysis, 20)
```

```{r}
ggplot(df_analysis, aes(x = customer_lifetime_value)) +
  geom_histogram(bins=40, fill="steelblue", color="white") +
  labs(title="Distribution of Customer Lifetime Value",
       x="Customer Lifetime Value",
       y="Frequency")
```

```{r}
cor_matrix <- cor(df_analysis)

melt(cor_matrix) %>%
  ggplot(aes(Var1, Var2, fill=value)) +
  geom_tile() +
  scale_fill_gradient2(low="blue", mid="white", high="purple") +
  theme(axis.text.x = element_text(angle=90)) +
  labs(title="Correlation Heatmap")
```

Above heatmap, contrast to linear regression models, decision tree algorithms are generally robust to multicollinearity among predictor variables. Since decision trees perform recursive data partitioning rather than estimating regression coefficients, highly correlated variables do not significantly distort model estimation. The features remain the same.

## 2.7 Decision Tree Analysis

### 2.7.1 Basic Decision Tree Model

```{r}
df_analysis <- df_analysis %>%
  mutate(clv_billion = customer_lifetime_value / 1e9)
```

```{r}
anova.model <- function(min_split, complexity_parameter, max_depth) {
  rpart(clv_billion ~ ., 
        data = df_analysis, 
        method = "anova", 
        control = rpart.control(minsplit = min_split, 
                                cp = complexity_parameter, 
                                maxdepth = max_depth))}

fit_tree <- anova.model(50, 0.01, 4)
```

### **2.7.2 Visualising the Decision Tree Model**

```{r}
visTree(fit_tree)
```

Fit the regression tree model.

```{r}
rpart.plot(fit_tree,
           type = 2,
           extra = 101,
           box.palette = "Blues",
           fallen.leaves = TRUE)

title("Pruned Decision Tree for Customer Lifetime Value (Billions COP)",
      line = 3,
      cex.main = 1.3)
```

### **2.7.3** Tuning of Hyperparameters

```{r}
printcp(fit_tree)
```

### **2.7.4** Pruned Tree

```{r}
bestcp <- fit_tree$cptable[which.min(fit_tree$cptable[,"xerror"]),"CP"]
```

```{r}
pruned_tree <- prune(fit_tree,cp=bestcp)
```

### **2.7.5** Visualizing Prune Tree

```{r}
visTree(pruned_tree)
```

```{r}
rpart.plot(pruned_tree,
           type = 2,
           extra = 101,
           box.palette = "Blues",
           fallen.leaves = TRUE)

title("Pruned Decision Tree for Customer Lifetime Value (Billions COP)",
      line = 3,
      cex.main = 1.3)
```

**Analysis:**

The pruning process selected a complexity parameter (CP = 0.01) that minimized the cross-validated error. Since this value corresponds to the largest tree generated during training, pruning did not remove any branches. As a result, the pruned tree remains identical to the original decision tree, indicating that the existing tree structure already provides the optimal balance between model complexity and predictive performance.

# 2.8 Random Forest

### 2.8.1 **Splitting of data set into train vs. test data**

```{r}
set.seed(1234)

trainIndex <- createDataPartition(df_analysis$customer_lifetime_value,
                                  p = 0.8,
                                  list = FALSE,
                                  times = 1)

df_train <- df_analysis[trainIndex, ]
df_test  <- df_analysis[-trainIndex, ]
```

### 2.8.2 Hyperparameter Tuning and Training of Model

The model is trained to predict **customer lifetime value** based on customer behavioural, transaction and satisfaction-related variables.

```{r}
trctrl <- trainControl(method = "none")

cvControl <- trainControl(method = "cv",
                          number = 10)

repeatcvControl <- trainControl(method = "repeatedcv",
                                number = 10,
                                repeats = 10)

rf_model <- train(customer_lifetime_value ~ .,
                  data = df_train,
                  method = "ranger",
                  trControl = repeatcvControl,
                  num.trees = 50,
                  importance = "impurity",
                  tuneGrid = data.frame(
                    mtry = floor(sqrt(ncol(df_train) - 1)),
                    min.node.size = 5,
                    splitrule = "variance"
                  ))
```

### Interpretation

The Random Forest model is designed to capture nonlinear relationships and interactions among the predictors without requiring strong distributional assumptions. Compared to a single decision tree, it generally produces more stable and accurate predictions by aggregating multiple trees trained on bootstrap samples.

### 2.8.3 Visualizing predicted vs. observed responses

The trained model is then applied to the test data to generate predicted customer lifetime values. Two plots are produced:

1.  Predicted vs. Actual scatterplot\
    to assess how closely the predictions align with the observed values.

2.  Residual plot\
    to examine the distribution of prediction errors and identify any systematic bias.

```{r}
#| eval: false
df_test$fit_forest <- predict(rf_model, df_test)

rf_scatter <- ggplot() +
  geom_point(aes(x = df_test$customer_lifetime_value,
                 y = df_test$fit_forest)) +
  labs(x = "Actual Customer Lifetime Value",
       y = "Predicted Customer Lifetime Value",
       title = paste0("R-squared: ",
                      round(cor(df_test$customer_lifetime_value,
                                df_test$fit_forest)^2, 2))) +
  theme(axis.text = element_text(size = 7),
        axis.title = element_text(size = 9),
        title = element_text(size = 10))

rf_residuals <- ggplot() +
  geom_point(aes(x = df_test$customer_lifetime_value,
                 y = (df_test$fit_forest - df_test$customer_lifetime_value)),
             col = "blue3") +
  labs(y = "Residuals (Predicted - Actual)",
       x = "Actual Customer Lifetime Value") +
  geom_hline(yintercept = 0,
             col = "red4",
             linetype = "dashed",
             linewidth = 0.7) +
  theme(axis.text = element_text(size = 7),
        axis.title = element_text(size = 9))

p <- rf_scatter + rf_residuals +
  plot_annotation(
    title = "Scatterplot of predicted vs. actual customer lifetime value",
    theme = theme(plot.title = element_text(size = 16))
  )

p
```

**Interpretation**

Chart(Left): The predicted versus actual scatterplot shows a strong linear relationship between predicted and observed customer lifetime values. Most points lie close to the diagonal trend, indicating that the Random Forest model is able to generate predictions that closely match the actual values. The model achieves an R² value of approximately 0.97, suggesting that it explains about 97% of the variance in customer lifetime value. This indicates strong predictive performance and suggests that the selected behavioural and transactional variables provide substantial information for estimating customer value.

Chart(Right): The residual plot illustrates the difference between predicted and actual customer lifetime values. Most residuals are distributed closely around zero, indicating that the model does not exhibit systematic bias in its predictions. The random dispersion of residuals suggests that the Random Forest model captures the underlying structure of the data effectively. However, a small number of larger residuals are observed for customers with extremely high lifetime values, indicating that the model may slightly underpredict some high-value customers.

### 2.8.4 Visualising Variable Importance (top 20)

```{r}
vip <- varImp(rf_model)
vip
plot(vip, top = 20)
```

### Interpretation

The variable importance chart identifies the most influential predictors of customer lifetime value. For this fintech dataset, transaction-related variables such as `total_tx_volume`, `tx_count`, `avg_tx_value`, or closely related behavioural indicators are expected to rank highly, since customers with larger and more frequent transaction activity tend to generate higher long-term value. This provides actionable business insight by highlighting which customer behaviours are most strongly associated with higher profitability.

### **2.8.5 Visualizing of R-Squred Value vs. The Number of Trees**

```{r}
#| eval: false
tree_range <- 5:150

rsquared_trees <- c()

for (i in tree_range) {

  rf_model_temp <- train(customer_lifetime_value ~ .,
                         data = df_train,
                         method = "ranger",
                         trControl = trctrl,
                         num.trees = i,
                         importance = "impurity",
                         tuneGrid = data.frame(
                           mtry = floor(sqrt(ncol(df_train) - 1)),
                           min.node.size = 5,
                           splitrule = "variance"
                         ))

  pred_temp <- predict(rf_model_temp, df_test)

  rsquared_trees <- append(rsquared_trees,
                           cor(df_test$customer_lifetime_value, pred_temp)^2)
}

rsquared_plot <- data.frame(tree_range, rsquared_trees)

ggplot(df = rsquared_plot) +
  geom_point(aes(x = tree_range, y = rsquared_trees)) +
  labs(x = "Number of trees",
       y = "R-squared values",
       title = "R-squared vs. Number of Trees Plot")
```

**Interpretation**

The R² versus number of trees plot illustrates how the predictive performance of the Random Forest model evolves as additional trees are added to the ensemble. Initially, when the forest contains only a small number of trees, the R² value is relatively low due to model instability and limited representation of the underlying data structure. As the number of trees increases, model performance improves rapidly and stabilises after approximately 50 trees. Beyond this point, the R² value fluctuates slightly around 0.96–0.97, indicating that further increases in tree count yield minimal improvement. This suggests that a moderate number of trees is sufficient to achieve strong predictive performance while maintaining computational efficiency.

```{r}
#| eval: false
var_imp <- varImp(rf_model_temp)

plot(var_imp, top = 10, main = "Top 10 Important Variables for CLV Prediction")
```

**Interpretation**

The variable importance plot indicates that transaction-related variables are the most influential predictors of Customer Lifetime Value (CLV) in the random forest model. In particular, total_tx_volume is the most important feature, suggesting that customers who generate higher overall transaction amounts tend to contribute significantly more lifetime value to the fintech platform. This is followed by tx_count and avg_tx_value, indicating that both the frequency and average size of transactions play a key role in determining customer value. Other variables such as average_transaction_value, credit_utilization_ratio, and monthly_transaction_count have moderate influence, reflecting the impact of customer spending behavior and financial activity. In contrast, variables such as churn_probability, weekend_transaction_ratio, avg_daily_transactions, and nps_score contribute relatively less to the prediction. Overall, the results suggest that customers who transact more frequently and generate higher transaction volumes tend to have greater lifetime value, highlighting the importance of transaction behavior in driving long-term customer profitability.

## Reference

1.  <https://ggplot2.tidyverse.org/>
2.  <https://cran.r-project.org/web/packages/ggparty/vignettes/ggparty-graphic-partying.html>
3.  <https://cran.r-project.org/web/packages/randomForestExplainer/randomForestExplainer.pdf>
