---
title: "Hands-on_Exercise1"
author: "Ho Tung"
date: "January 10, 2026"
date-modified: "last-modified"

format:
  html:
    toc-depth: 4
    toc-expand: true

execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# 1.1 Overview

Basic principles and essential components of ggplot2

## **1.2 Getting started**

```{r}
pacman::p_load(tidyverse)
```

### **1.2.2 Importing data**

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

## **1.3 Introducing ggplot**

### 1.3.1 R Graphics VS ggplot 
::: {.panel-tabset}

#### R Graphics

```{r}
hist(exam_data$MATHS)
```

#### ggplot

```{r}
ggplot(data=exam_data, aes(x = MATHS)) +
  geom_histogram(bins=10, 
                 boundary = 100,
                 color="black", 
                 fill="grey") +
  ggtitle("Distribution of Maths scores")
```
:::


## **1.4 Grammar of Graphics**

### **1.4.1 A Layered Grammar of Graphics**

## **1.5 Essential Grammatical Elements in ggplot2: data**

```{r}
ggplot(data=exam_data)
```

## **1.6 Essential Grammatical Elements in ggplot2: [Aesthetic mappings](https://ggplot2.tidyverse.org/articles/ggplot2-specs.html)**

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS))
```

## **1.7 Essential Grammatical Elements in ggplot2: geom 1.7.1 Geometric Objects: geom_bar**

```{r}
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar()
```

### **1.7.2 Geometric Objects: geom_dotplot**

```{r}
ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot(dotsize = 0.5)
```

```{r}
ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot(binwidth=2.5,         
               dotsize = 0.5) +      
  scale_y_continuous(NULL,           
                     breaks = NULL)  
```

### **1.7.3 Geometric Objects: `geom_histogram()`**

```{r}
ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_histogram()       
```

### **1.7.4 Modifying a geometric object by changing `geom()`**

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS)) +
  geom_histogram(bins=20,            
                 color="black",      
                 fill="light blue")  
```

### **1.7.5 Modifying a geometric object by changing *aes()***

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           fill = GENDER)) +
  geom_histogram(bins=20, 
                 color="grey30")
```

### **1.7.6 Geometric Objects: geom-density()**

```{r}
ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_density()           
```

```{r}
ggplot(data=exam_data, 
       aes(x = MATHS, 
           colour = GENDER)) +
  geom_density()
```

### **1.7.7 Geometric Objects: geom_boxplot**

```{r}
ggplot(data=exam_data,
       aes(y = MATHS,       
           x= GENDER)) +    
  geom_boxplot()             
```

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, 
           x= GENDER)) +
  geom_boxplot(notch=TRUE)
```

### **1.7.8 Geometric Objects: geom_violin**

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, 
           x= GENDER)) +
  geom_violin()
```

### **1.7.9 Geometric Objects: geom_point()**

[`geom_point()`](https://ggplot2.tidyverse.org/reference/geom_point.html)

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point()            
```

### **1.7.10 *geom* objects can be combined**

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, 
           x= GENDER)) +
  geom_boxplot() +                    
  geom_point(position="jitter", 
             linewidth = 0.5)        
```

## **1.8 Essential Grammatical Elements in ggplot2: stat**

### **1.8.1 Working with `stat()`**

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, x= GENDER)) +
  geom_boxplot()
```

### **1.8.2 Working with stat - the *stat_summary()* method**

### [`stat_summary()`](https://ggplot2.tidyverse.org/reference/stat_summary.html) function

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, x= GENDER)) +
  geom_boxplot()
```

### **1.8.2 Working with stat - the *stat_summary()* method** [`stat_summary()`](https://ggplot2.tidyverse.org/reference/stat_summary.html) 

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, x= GENDER)) +
  geom_boxplot() +
  stat_summary(geom = "point",       
               fun = "mean",         
               colour ="red",        
               linewidth=4)               
```

### **1.8.3 Working with stat - the `geom()` method**

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, x= GENDER)) +
  geom_boxplot() +
  geom_point(stat="summary",        
             fun="mean",           
             colour="red",          
             linewidth=4)          
```

### **1.8.4 Adding a best fit curve on a scatterplot?**

[`geom_smooth()`](https://ggplot2.tidyverse.org/reference/geom_smooth.html) 

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, y=ENGLISH)) +
  geom_point() +
  geom_smooth(linewidth=0.5)
```

```{r}

```

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              linewidth=0.5)
```

## **1.9 Essential Grammatical Elements in ggplot2: Facets**

[`facet_grid()`](https://ggplot2.tidyverse.org/reference/facet_grid.html) and [`facet_wrap`](https://ggplot2.tidyverse.org/reference/facet_wrap.html)

### **1.9.1 Working with `facet_wrap()`**

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS)) +
  geom_histogram(bins=20) +
    facet_wrap(~ CLASS)
```

### **1.9.2 `facet_grid()` function**

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS)) +
  geom_histogram(bins=20) +
    facet_grid(~ CLASS)
```

## **1.10 Essential Grammatical Elements in ggplot2: Coordinates**

\- \[\`coord_cartesian()\`\](https://ggplot2.tidyverse.org/reference/coord_cartesian.html): the default cartesian coordinate systems, where you specify x and y values (e.g. allows you to zoom in or out). - \[\`coord_flip()\`\](https://ggplot2.tidyverse.org/reference/coord_flip.html): a cartesian system with the x and y flipped. - \[\`coord_fixed()\`\](https://ggplot2.tidyverse.org/reference/coord_fixed.html): a cartesian system with a "fixed" aspect ratio (e.g. 1.78 for a "widescreen" plot). - \[\`coord_quickmap()\`\](https://ggplot2.tidyverse.org/reference/coord_map.html): a coordinate system that approximates a good aspect ratio for maps.

\
**1.10.1 Working with Coordinate**

```{r}
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar()
```

```{r}
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar() +
  coord_flip()
```

### **1.10.2 Changing the y- and x-axis range**

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, linewidth=0.5)
```

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              linewidth=0.5) +  
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))
```

## **1.11 Essential Grammatical Elements in ggplot2: themes**

### **1.11.1 Working with theme**

```{r}
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar() +
  coord_flip() +
  theme_gray()
```

```{r}
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar() +
  coord_flip() +
  theme_classic()
```

```{r}
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar() +
  coord_flip() +
  theme_classic()
```

```{r}
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar() +
  coord_flip() +
  theme_classic()
```

## **1.12 Reference**

-   Hadley Wickham (2023) [**ggplot2: Elegant Graphics for Data Analysis**](https://ggplot2-book.org/). Online 3rd edition.

-   Winston Chang (2013) [**R Graphics Cookbook 2nd edition**](https://r-graphics.org/). Online version.

-   Healy, Kieran (2019) [**Data Visualization: A practical introduction**](https://socviz.co/). Online version

-   [Learning ggplot2 on Paper – Components](https://henrywang.nl/learning-ggplot2-on-paper-components/)

-   [Learning ggplot2 on Paper – Layer](https://henrywang.nl/learning-ggplot2-on-paper-layer/)

-   [Learning ggplot2 on Paper – Scale](https://henrywang.nl/tag/learning-ggplot2-on-paper/)
