---
title: "Hands-on_Exercise7"
author: "Chen Ho Tung"
date: "Feburary 23, 2026"
date-modified: "last-modified"

format:
  html:
    toc-depth: 4
    toc-expand: true

execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# **17  Visualising and Analysing Time-oriented Data**

## **17.1 Learning Outcome**

-   plotting a calender heatmap by using ggplot2 functions,

-   plotting a cycle plot by using ggplot2 function,

-   plotting a slopegraph

-   plotting a horizon chart

## **17.2 Getting Started**

## **17.3 Import**

```{r}
pacman::p_load(scales, viridis, lubridate, ggthemes,
               gridExtra, readxl, knitr, data.table,
               CGPfunctions, ggHoriPlot, tidyverse)
```

## **17.4 Plotting Calendar Heatmap**

### **17.4.1 The Data**

### **17.4.2 Importing the data**

```{r}
attacks <- read_csv("data/eventlog.csv")
```

### **17.4.3 Examining the data structure**

For example, *kable()* can be used to review the structure of the imported data frame.

```{r}
kable(head(attacks))
```

### **17.4.4 Data Preparation**

Step 1: Deriving *weekday* and *hour of day* fields

```{r}
make_hr_wkday <- function(ts, sc, tz) {
  real_times <- ymd_hms(ts, 
                        tz = tz[1], 
                        quiet = TRUE)
  dt <- data.table(source_country = sc,
                   wkday = weekdays(real_times),
                   hour = hour(real_times))
  return(dt)
}
```

Step 2: Deriving the attacks tibble data frame

```{r}
wkday_levels <- c('Saturday', 'Friday', 
                  'Thursday', 'Wednesday', 
                  'Tuesday', 'Monday', 
                  'Sunday')

attacks <- attacks %>%
  group_by(tz) %>%
  do(make_hr_wkday(.$timestamp, 
                   .$source_country, 
                   .$tz)) %>% 
  ungroup() %>% 
  mutate(wkday = factor(
    wkday, levels = wkday_levels),
    hour  = factor(
      hour, levels = 0:23))
```

Table below shows the tidy tibble table after processing.

```{r}
kable(head(attacks))
```

### **17.4.5 Building the Calendar Heatmaps**

```{r}
#| warning: false
grouped <- attacks %>% 
  count(wkday, hour) %>% 
  ungroup() %>%
  na.omit()

ggplot(grouped, 
       aes(hour, 
           wkday, 
           fill = n)) + 
geom_tile(color = "white", 
          size = 0.1) + 
theme_tufte(base_family = "Helvetica") + 
coord_equal() +
scale_fill_gradient(name = "# of attacks",
                    low = "sky blue", 
                    high = "dark blue") +
labs(x = NULL, 
     y = NULL, 
     title = "Attacks by weekday and time of day") +
theme(axis.ticks = element_blank(),
      plot.title = element_text(hjust = 0.5),
      legend.title = element_text(size = 8),
      legend.text = element_text(size = 6) )
```

### **17.4.6 Building Multiple Calendar Heatmaps**

**Challenge:** Building multiple heatmaps for the top four countries with the highest number of attacks.

### **17.4.7 Plotting Multiple Calendar Heatmaps**

Step 1: Deriving attack by country object

```{r}
attacks_by_country <- count(
  attacks, source_country) %>%
  mutate(percent = percent(n/sum(n))) %>%
  arrange(desc(n))
```

Step 2: Preparing the tidy data frame

```{r}
top4 <- attacks_by_country$source_country[1:4]
top4_attacks <- attacks %>%
  filter(source_country %in% top4) %>%
  count(source_country, wkday, hour) %>%
  ungroup() %>%
  mutate(source_country = factor(
    source_country, levels = top4)) %>%
  na.omit()
```

### **17.4.8 Plotting Multiple Calendar Heatmaps**

Step 3: Plotting the Multiple Calender Heatmap by using ggplot2 package.

```{r}
#| warning: false
ggplot(top4_attacks, 
       aes(hour, 
           wkday, 
           fill = n)) + 
  geom_tile(color = "white", 
          size = 0.1) + 
  theme_tufte(base_family = "Helvetica") + 
  coord_equal() +
  scale_fill_gradient(name = "# of attacks",
                    low = "sky blue", 
                    high = "dark blue") +
  facet_wrap(~source_country, ncol = 2) +
  labs(x = NULL, y = NULL, 
     title = "Attacks on top 4 countries by weekday and time of day") +
  theme(axis.ticks = element_blank(),
        axis.text.x = element_text(size = 7),
        plot.title = element_text(hjust = 0.5),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 6) )
```

## **17.5 Plotting Cycle Plot**

### **17.5.1 Step 1: Data Import**

For the purpose of this hands-on exercise, *arrivals_by_air.xlsx* will be used.

The code chunk below imports *arrivals_by_air.xlsx* by using `read_excel()` of **readxl** package and save it as a tibble data frame called *air*.

```{r}
air <- read_excel("data/arrivals_by_air.xlsx")
```

### **17.5.2 Step 2: Deriving month and year fields**

```{r}
air$month <- factor(month(air$`Month-Year`), 
                    levels=1:12, 
                    labels=month.abb, 
                    ordered=TRUE) 
air$year <- year(ymd(air$`Month-Year`))
```

### **17.5.3 Step 4: Extracting the target country**

Next, the code chunk below is use to extract data for the target country (i.e. Vietnam)

```{r}
Vietnam <- air %>% 
  select(`Vietnam`, 
         month, 
         year) %>%
  filter(year >= 2010)
```

### **17.5.4 Step 5: Computing year average arrivals by month**

```{r}
hline.data <- Vietnam %>% 
  group_by(month) %>%
  summarise(avgvalue = mean(`Vietnam`))
```

### **17.5.5 Srep 6: Plotting the cycle plot**

The code chunk below is used to plot the cycle plot as shown in Slide 12/23.

```{r}
#| warning: false
ggplot() + 
  geom_line(data=Vietnam,
            aes(x=year, 
                y=`Vietnam`, 
                group=month), 
            colour="black") +
  geom_hline(aes(yintercept=avgvalue), 
             data=hline.data, 
             linetype=6, 
             colour="red", 
             size=0.5) + 
  facet_grid(~month) +
  labs(axis.text.x = element_blank(),
       title = "Visitor arrivals from Vietnam by air, Jan 2010-Dec 2019") +
  xlab("") +
  ylab("No. of Visitors") +
  theme_tufte(base_family = "Helvetica")
```

## **17.6 Plotting Slopegraph**

In this section you will learn how to plot a [slopegraph](https://www.storytellingwithdata.com/blog/2020/7/27/what-is-a-slopegraph) by using R.

Before getting start, make sure that **CGPfunctions** has been installed and loaded onto R environment. Then, refer to [Using newggslopegraph](https://cran.r-project.org/web/packages/CGPfunctions/vignettes/Using-newggslopegraph.html) to learn more about the function. Lastly, read more about `newggslopegraph()` and its arguments by referring to this [link](https://www.rdocumentation.org/packages/CGPfunctions/versions/0.6.3/topics/newggslopegraph).

### **17.6.1 Step 1: Data Import**

Import the rice data set into R environment by using the code chunk below.

```{r}
rice <- read_csv("data/rice.csv")
```

### **17.6.2 Step 2: Plotting the slopegraph**

Next, code chunk below will be used to plot a basic slopegraph as shown below.

```{r}
rice %>% 
  mutate(Year = factor(Year)) %>%
  filter(Year %in% c(1961, 1980)) %>%
  newggslopegraph(Year, Yield, Country,
                Title = "Rice Yield of Top 11 Asian Counties",
                SubTitle = "1961-1980",
                Caption = "Prepared by: Dr. Kam Tin Seong")
```
